```{r}
rm(list=ls())
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(plotly))
suppressPackageStartupMessages(library(MASS))

df_alpha <- read_csv("../data/alpha_event_data.csv", show_col_types = FALSE)
df_beta <- read_csv("../data/beta_event_data.csv", show_col_types = FALSE)
df_gamma <- read_csv("../data/gamma_event_data.csv", show_col_types = FALSE)
band_names <- rep(c('alpha', 'beta', 'gamma'),
                  times=c(nrow(df_alpha), nrow(df_beta), nrow(df_gamma)))

df <- rbind(df_alpha, df_beta, df_gamma)
names(df) <- tolower(str_replace_all(names(df_alpha), ' ', '_')) # clean names
df['freq_band'] <- band_names
df['position'] <- as.character(df$position)
df['layout'] <- as.character(df$layout)
```

```{r}
median_reaction_time <- median(df$reaction_time, na.rm = TRUE)
median_reach_time <- median(df$reach_time, na.rm = TRUE)

# Calculate the statistics and add a binary variable column based on the median reaction_time
stats_df <- df %>%
  group_by(trial, freq_band) %>%
  summarize(
    mean_freq_span = mean(frequency_span, na.rm = TRUE),
    std_freq_span = sd(frequency_span, na.rm = TRUE),
    mean_event_duration = mean(event_duration, na.rm = TRUE),
    std_event_duration = sd(event_duration, na.rm = TRUE),
    mean_peak_time = mean(peak_time, na.rm = TRUE),
    std_peak_time = sd(peak_time, na.rm = TRUE),
    mean_peak_power = mean(peak_power, na.rm = TRUE),
    std_peak_power = sd(peak_power, na.rm = TRUE),
    count = n(),
    .groups = 'drop'
  ) 

# Add reaction_time, reach_time, and reaction_time_binary columns
summary_df <- df %>%
  group_by(trial) %>%
  summarize(
    reaction_time = first(reaction_time),
    reach_time = first(reach_time),
    reaction_time_binary = ifelse(first(reaction_time) > median_reaction_time, 1, 0),
    reach_time_binary = ifelse(first(reach_time) > median_reach_time, 1, 0),
    .groups = 'drop'
  )

# Pivot the statistics dataframe from long to wide format
output_df <- stats_df %>%
  pivot_wider(
    id_cols = trial,
    names_from = freq_band,
    values_from = c(mean_freq_span, std_freq_span, mean_event_duration, std_event_duration, mean_peak_time, std_peak_time, mean_peak_power, std_peak_power, count),
    names_sep = "_"
  ) %>%
  left_join(summary_df, by = "trial")

output_df
```

```{r}
output_df <- na.omit(output_df)

# linear regression (reaction time, full model)
lm_reaction <- lm(reaction_time ~ . - reach_time - reaction_time_binary - reach_time_binary, data = output_df)

# backward and forward stepwise regression
step_reaction <- stepAIC(lm_reaction, direction = "both")
summary(step_reaction)$coef
```

```{r}
# linear regression (reach time, full model)
lm_reach <- lm(reach_time ~ . - reaction_time - reaction_time_binary - reach_time_binary, data = output_df)

# backward and forward stepwise regression
step_reach <- stepAIC(lm_reach, direction = "both")
summary(step_reach)$coef
```

```{r}
# logistic regression (reaction time)
logit_reaction <- glm(reaction_time_binary ~ ., family = binomial(link = "logit"), data = output_df)
summary(logit_reaction)$coef
```

```{r}
# logistic regression (reach time)
logit_reach <- glm(reach_time_binary ~ ., family = binomial(link = "logit"), data = output_df)
summary(logit_reach)$coef
```

```{r}
# instantiates matrices to contain the beta coefficients for all 1000 samples
boot <- 1000
beta1_boot <- matrix(NA, boot, 1)
beta2_boot <- matrix(NA, boot, 1)
beta3_boot <- matrix(NA, boot, 1)
beta4_boot <- matrix(NA, boot, 1)
beta5_boot <- matrix(NA, boot, 1)
beta6_boot <- matrix(NA, boot, 1)
beta7_boot <- matrix(NA, boot, 1)

# performs the bootstrap iteratively to compute all beta coefficients
for (i in 1:boot) {
  s_boot <- sample(c(1:dim(output_df)[1]), dim(output_df)[1], replace = TRUE)
  data_boot <- output_df[s_boot,]
  fit_boot <- lm(reaction_time ~ mean_freq_span_gamma + mean_event_duration_gamma 
                 + std_event_duration_gamma + std_peak_time_beta + std_peak_power_beta 
                 + count_beta + count_gamma, data = data_boot)
  beta1_boot[i,] <- fit_boot$coef[2]
  beta2_boot[i,] <- fit_boot$coef[3]
  beta3_boot[i,] <- fit_boot$coef[4]
  beta4_boot[i,] <- fit_boot$coef[5]
  beta5_boot[i,] <- fit_boot$coef[6]
  beta6_boot[i,] <- fit_boot$coef[7]
  beta7_boot[i,] <- fit_boot$coef[8]
}

# prepares a dataframe with both bootstrap and backward procedure coefficients
df_bootstrap_lm_reaction <- data.frame(beta = rep(c("Beta1: Mean_Freq_Span_Gamma", "Beta2: Mean_Event_Duration_Gamma", 
                                                    "Beta3: STD_Event_Duration_Gamma", "Beta4: STD_Peak_Time_Beta", 
                                                    "Beta5: STD_Peak_Power_Beta", "Beta6: Count_Beta", 
                                                    "Beta7: Count_Gamma"), each = boot),
                                       estimate = c(beta1_boot, beta2_boot, beta3_boot, beta4_boot, beta5_boot, 
                                                    beta6_boot, beta7_boot))

# creates a boxplot of the bootstrap results for each beta
plot_bootstrap_lm_reaction <- ggplot(df_bootstrap_lm_reaction, aes(x = beta, y = estimate, fill = beta)) +
  xlab("Beta") +
  ylab("Estimate") +
  ggtitle("Bootstrap Results") +
  geom_boxplot(na.rm = TRUE) +
  labs(fill = "Beta") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
plot_bootstrap_lm_reaction

# creates a histogram of the bootstrap results for each beta
plot_bootstrap_lm_reaction <- ggplot(df_bootstrap_lm_reaction, aes(x = estimate)) +
  ggtitle("Bootstrap Results") +
  xlab("Estimate") +
  ylab("Count") +
  geom_histogram(color = "gray15", fill = "gray90", binwidth = 5) +
  facet_wrap(~beta) +
  theme_bw()
plot_bootstrap_lm_reaction
```

The boxplots indicate that the bootstrapped samples, with the exception of the "event duration" variable, have very narrow distributions, and the histograms illustrate how the estimates of the proportion of variance explained are clustered tightly in a small range of values. The bootstrap allows us to assess the variability of the population estimate by repeatedly sampling observations from the original dataset. This approach simulates the process of obtaining independent datasets from the population, enabling us to generalize our results to the population. Therefore, since the range of every boxplot is very small, we have increased confidence in the accuracy, stability, and generalizability of the proportion estimates generated by the linear regression model.